### 2022-07-22 main/chapter_2ljets.tex
expand zjets alt section

\subsection{Sample comparisons}
\label{sec:2ljets_sample_comparisons}
The Standard Model is a distillation of past data, and all of our simulations
of it have been tuned to those data, in both their treatment of fundamental
physics and of experimental hardware.%
\footnote{%
Amusingly, tuning to data specific to each experiment means that each
experiment can simulate different fundamental physics~\cite{
herwig2000parameter,
Altarelli:473529
}, not only different detector responses.
All Standard Models are not the same.
}
Each tune yields a point estimate that does not incorporate remaining
uncertainties, so those uncertainties must be quantified by other means.
Some uncertainties are included as weights, particularly those for
uncertainties in the
PDFs (parton distribution functions) and phyiscal scales
(renormalization and factorization scales, $\mu_R$ and $\mu_F$).
Our inclusion of these weighty variations through standard prescriptions is
described in Section~\ref{sec:2ljets_theory_variations}.

@@

Clearly, it could happen that two algorithms agree in their tune despite
neither being confident --- as an illustration, imagine a wide error bar and
two post-fit results right in its centre.
This effect would yield an underestimated uncertainty, so does not invalidate
the comparison approach.
On the contrary, it highlights that it is a bare minimum and that larger
uncertainties may well be reasonable given our imprecise constraints on the
Standard Model.

Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt}
Figure~\ref{fig:2ljets_low_crz_pre_post}
Figure~\ref{fig:2ljets_vrlow_alt}
Figure~\ref{fig:2ljets_low_vrlow_pre_post}

For implementations in the HistFactory model, these variations reduce to
the relative scale factors presented in Table~\ref{tab:2ljets_zjets_alt}.
Note that up and down variations of the sample comparison parameter do not
mean up and down variations of bin yields, but to the coupled variations
in directions given by the signs of the relative variations given in the
various regions.

\begin{table}[tp]
\centering
\begin{tabular}{lcc}
& Central
& Relative variations
\\[0.5em]
CR-DY
& $59.64$
& $1-0.1+0.1$
\\[0.2em]
\underline{CR-Z}
& $101.95$
& $1+0.2-0.2$
\\[0.2em]
CR-tt
& $38.07$
& $1+0.5-0.5$
\\[0.2em]
CR-VZ
& $5.03$
& $1+1.3-1.0$
\\[0.5em]
VR-Low
& $24.19$
& $1-0.3+0.3$
\\[0.2em]
VR-Low-2
& $17.05$
& $1+0.2-0.2$
\\[0.5em]
\underline{SR-Low-a}
& $14.30$
& $1-0.2+0.2$
\\[0.2em]
\underline{SR-Low-b}
& $4.17$
& $1-0.3+0.3$
\\[0.2em]
SR-Low-2
& $1.51$
& $1+1.5-1.0$
\\[0.2em]
\end{tabular}
\caption[%
Alternative $\zjets$ variations in selected regions
]{%
Alternative $\zjets$ variations in selected regions that have significant
$\zjets$ contributions.
In other regions the $\zjets$ sample and its variations are negligible.
The `Central' column shows the central $\zjets$ yield from the main sample.
The other column shows the relative scale factors applied to this central yield
for up and down variations.
These are evaluated from the central and alternative samples as described in
the text.
Note how up and down variations can move in different directions, and these
different movements are coupled in the background model.
}
\label{tab:2ljets_zjets_alt}
\end{table}

>>>>

\subsection{Sample comparisons}
\label{sec:2ljets_sample_comparisons}
The Standard Model is a distillation of past data, and all of our simulations
of it have been tuned to those data, in both their treatment of fundamental
physics and of experimental hardware.%
\footnote{%
Amusingly, tuning to data specific to each experiment means that each
experiment can simulate different fundamental physics~\cite{
herwig2000parameter,
Altarelli:473529
}, not only different detector responses.
All Standard Models are not the same.
}
Each tune yields a point estimate that does not incorporate remaining
uncertainties, so those uncertainties must be quantified by other means.
Some uncertainties are included as weights, particularly those for
uncertainties in the
PDFs (parton distribution functions) and physical scales
(renormalization and factorization scales, $\mu_R$ and $\mu_F$).
Our inclusion of these weighty variations through standard prescriptions is
described in Section~\ref{sec:2ljets_theory_variations}.

@@

Clearly, it could happen that two algorithms agree in their tune despite
neither being confident.
To illustrate this point, imagine a wide error bar and
two post-fit results right in its centre;
both have achieved the best fit, but both also could allow large variations.
This effect would yield an underestimated uncertainty, so does not invalidate
the comparison approach but highlights that it is a bare minimum and that
larger uncertainties may well be reasonable given our imprecise constraints on
the Standard Model.

As mentioned in Section~\ref{sec:2ljets_low}, the $\zjets$ sample comparison
is hugely important to our results in SR-Low.
Its central yields are displayed in Figure~\ref{fig:2ljets_low_sr_or_cr_region}
in the union of CR-Z and SR-Low, which form a continuous region in $\mjj$.
Our central $\zjets$ samples use \sherpa~2.2.1~\cite{
Bothmann:2019yzt,
ATL-PHYS-PUB-2016-003
}
at next-to-leading order for up to $4$ partonic jets, and leading order for
$5$ partonic jets.
In Figure~\ref{fig:2ljets_low_sr_or_cr_region_alt} we compare this to our
alternative sample that simulates its hard scatter
with \madgraph\ (MG5\_aMC@NLO~2.2.3~\cite{Alwall:2014hca}) at leading order with
up to four partonic jets, plugged into \pythia~8~\cite{Sjostrand:2014zea} for
showering and hadronization.
These plots show statistical error bars only, and the larger statistical error
bars on the alternative sample are due to it comprising around one sixth as
many simulated events as the central sample.

For their implementation in the background model in HistFactory, these
variations reduce to the relative scale factors that are presented in
Table~\ref{tab:2ljets_zjets_alt}.
Note that up and down variations of the sample comparison parameter do not
mean up and down variations of bin yields, but to the coupled variations
in directions given by the signs of the relative variations given in the
various regions.
These variations are evaluated according to a standard recipe prescribed by
the ATLAS SUSY group~\cite{atlas_twiki_susy_theory_comparison} that says to
\begin{itemize}
\item evaluate the difference between the two samples $\delta_{12}$.
\item evaluate the statistical uncertainty on the alternative sample $\sigma_2$.
\item combine these components in quadrature, so
$\sigma_{12} = \sqrt{\delta_{12}^2 + \sigma_2^2}$, and
\item if the deviation is statistically significant, then preserve the direction
of the uncertainty, so $\sigma = \mathrm{sgn}\sigma_{12}$.
\end{itemize}
An `if A then B' statement does not define what to do if A is false, so
this recipe does not define what sign to give $\sigma$ if the deviation is not
statistically significant.
Since all variations have direction
--- they are either positive or negative or zero ---
we are unable to give no direction to any deviation, and we choose to
we use the same, signed, definition of $\sigma$ for all regions.

As seen from Table~\ref{tab:2ljets_zjets_alt}, the statistical uncertainty
$\sigma_2$ is a substantial contribution to these variations;
the statistical uncertainty in the alternative sample does significantly
expand this variation.
Indeed, it is the larger contribution in SR-Low-a and SR-Low-b, which are the
regions most sensitive to $\zjets$ variations.
Our unprescribed choice to match their directions therefore matters.
It may have been reasonable to treat them differently, perhaps by adding
independent parameters in the separate regions.
I believe such a change would have improved the plausibility of our background
modelling, and I regret not finding the capacity to push it at the time.

This SUSY recipe contradicts the advice of the ATLAS Physics Modelling Group
(PMG), who say to quote only the absolute value of the two
estimates~\cite{atlas_twiki_pmg_theory_comparison} in for two-sample
comparisons.
Since our search is in the SUSY group we follow their instructions.

Figure~\ref{fig:2ljets_low_crz_pre_post}
Figure~\ref{fig:2ljets_vrlow_alt}
Figure~\ref{fig:2ljets_low_vrlow_pre_post}

% in branch hackmg5
% python get_ewk_Zjets_alt.py
%                 CRDY,     2.30,     5.97
%                  CRZ,    16.75,     6.68
%                 CRZZ,     6.35,     1.61
%                 CRtt,    18.06,     4.29
%           SRHigh16_1,     0.02,     0.00
%           SRHigh16_2,     0.01,     0.00
%              SRHigh4,     0.02,     0.00
%            SRHigh8_1,     0.02,     0.00
%            SRHigh8_2,     0.02,     0.00
%              SRInt_1,     0.46,     0.76
%              SRInt_2,     0.09,     0.20
%               SRLow2,     1.63,     1.69
%              SRLow_1,     2.20,     2.36
%              SRLow_2,     0.45,     1.02
%         SROffShell_1,     0.12,     0.30
%         SROffShell_2,     0.95,     0.95
%               SRllbb,     0.03,     0.00
%               VRHigh,     0.13,     0.14
%              VRHigh4,     0.49,     0.39
%              VRHighR,     0.26,     0.34
%                VRInt,     0.32,     0.23
%                VRLow,     7.34,     3.24
%               VRLow2,     1.71,     3.21
%           VROffShell,     0.09,     1.06
%               VRllbb,     0.31,     0.36
\begin{table}[tp]
\centering
\begin{tabular}{lcccc}
& Central
& Relative variations
& $|\delta_{12}|$
& $\sigma_2$
\\[0.5em]
CR-DY
& $59.64$
& $1-0.1+0.1$
& $2.3$
& $6.0$
\\[0.2em]
\underline{CR-Z}
& $101.95$
& $1+0.2-0.2$
& $16.8$
& $6.7$
\\[0.2em]
CR-tt
& $38.07$
& $1+0.5-0.5$
& $18.1$
& $4.29$
\\[0.2em]
CR-VZ
& $5.03$
& $1+1.3-1.0$
& $6.4$
& $1.6$
\\[0.5em]
VR-Low
& $24.19$
& $1-0.3+0.3$
& $7.3$
& $3.2$
\\[0.2em]
VR-Low-2
& $17.05$
& $1+0.2-0.2$
& $1.7$
7 $3.2$
\\[0.5em]
\underline{SR-Low-a}
& $14.30$
& $1-0.2+0.2$
& $2.2$
& $2.4$
\\[0.2em]
\underline{SR-Low-b}
& $4.17$
& $1-0.3+0.3$
& $0.5$
& $0.8$
\\[0.2em]
SR-Low-2
& $1.51$
& $1+1.5-1.0$
& $1.6$
& $1.7$
\\[0.2em]
\end{tabular}
\caption[%
Alternative $\zjets$ variations in selected regions
]{%
Alternative $\zjets$ variations in selected regions that have significant
$\zjets$ contributions.
In other regions the $\zjets$ sample and its variations are negligible.
The `Central' column shows the central $\zjets$ yield from the main sample.
The other column shows the relative scale factors applied to this central yield
for up and down variations.
These are evaluated from the central and alternative samples as described in
the text of Section~\ref{sec:2ljets_sample_comparisons};
$\delta_{12}$ is the difference of the alternative from the central sample
and $\sigma_2$ is the statistical uncertainty on the alternative sample.
Note that the up and down variations can move in different directions, and
these different movements are coupled in the background model.
}
\label{tab:2ljets_zjets_alt}
\end{table}
